# Results {#sec-results}

In this chapter, the results of our analysis will be presented. In a first step, we will conduct a descriptive and exploratory analysis in @sec-exploratory-analysis of the data used for modelling. In @sec-model-evaluation, we will present and compare the results of the different models. Finally in @sec-final-model, we will provide the results of the final model and the sensitivity analysis of the input features.

```{r}
#| label: load-libs
#| include: false

library(tidyverse)
library(tidymodels)
library(doParallel)
library(skimr)
library(plotly)
library(sf)
library(agua)
library(gt)
library(rules)
library(baguette)
library(ggridges)
library(viridis)
library(hrbrthemes)
library(finetune)
library(ggrepel)
library(vip)
library(shapviz)
library(DALEXtra)
```

```{r}
#| label: load-data
#| include: false

energy_modelling <- read_rds("data/energy_modelling.rds")
```

## Exploratory Analysis {#sec-exploratory-analysis}

To gain a first overview of the data that we use for modelling, we here present the summary statistics and visual exploration of the data.

### Summary statistics

As outlined in @sec-data-prep, we applied several filters to the initial raw data. Thus, the final dataset for modelling contains `r nrow(energy_modelling)` observations (buildings). As we consider data from the years 2016, 2018 and 2020, one building can be present in the dataset three times at maximum. @tbl-obs-per-year shows that the number of gas-heated buildings is slightly decreasing between 2016 and 2020. Also, we observe a decrease of the mean HEPI and the total HEC between 2016 and 2020.

```{r}
#| label: tbl-obs-per-year
#| tbl-cap: "Number of buildings per survey year."
#| tbl-cap-location: bottom
#| warning: false

obs_per_year <- energy_modelling |>
  dplyr::group_by(survey_year) |>
  dplyr::rename("Survey Year" = survey_year) |>
  dplyr::summarise(Count = n(),
                   "Mean HEPI (kWh/m2/year)" = mean(hepi),
                   "Sum HEC (GWh/year)" = sum(hec)/1000) |>
  knitr::kable(booktabs = TRUE, digits = 1)

obs_per_year
```

```{r}
#| label: tbl-summary-stats
#| tbl-cap: "Summary statistics of modelling input data."
#| tbl-cap-location: bottom
#| eval: false

summary_stats_numeric <- energy_modelling |>
  skim() |>
  yank("numeric")

summary_stats_numeric
```

### Retrofitting

```{r}
#| label: indicators-retrofit
#| echo: false

#> Calculations of numbers and indicators for easier use as inline code in text.

n_retrofit <- energy_modelling |> filter(retrofitted == TRUE) |> nrow()

n_retrofit_m01 <- energy_modelling |> 
  filter(retrofitted == TRUE,
         retrofit_investment_costs > 0) |> 
  nrow()

perc_retrofit <- round((n_retrofit / nrow(energy_modelling))*100,digits = 2)
```

In the total dataset, we observe `r n_retrofit` building that could be matched with a retrofit measure. This represents `r perc_retrofit` % of all the buildings in the data that we use for modelling.

The retrofit data considers retrofit measures that have been financially supported by the cantonal Förderprogramm (Baselbieter Energiepaket) since 2009.

```{r}
#| label: fig-retrofit-year
#| fig-cap: "Distribution of retrofit year."

retrofit_per_year <- energy_modelling |>
  dplyr::filter(retrofitted == TRUE) |>
  ggplot((aes(retrofit_year_completion))) +
           geom_bar(aes(fill = retrofit_project_name)) +
  scale_x_continuous(breaks=seq(2009, 2019, 1))
  
  
retrofit_per_year
```

A first explorative way of analyzing the effect of retrofit measures consists in comparing the mean HEPI of retrofitted building and non-retrofitted buildings. @tbl-retrofit shows the mean HEPI per survey year and building class. For all years and building classes, the mean HEPI of retrofitted buildings is lower. Furthermore, a more important difference in average HEPI between retrofitted and non-retrofitted buildings can be observed for multi-family buildings (buildings with two or more apartments) over the course of all three years. This could potentially indicate that retrofit measures are more effective in terms of energy consumption reduction for multi-family houses in comparison to single-family houses.

```{r}
#| label: tbl-retrofit
#| tbl-cap: "Mean HEPI of retrofitted versus non-retrofitted."
#| tbl-cap-location: bottom
#| eval: false

retrofit_mean_hepi <- energy_modelling |>
  dplyr::group_by(survey_year, building_class, retrofitted) |>
  dplyr::rename("Survey Year" = survey_year) |>
  dplyr::summarise("Mean HEPI (kWh/m2/year)" = mean(hepi)) |>
  mutate(building_class = factor(building_class, 
                                  levels = c("1110", "1121", "1122"), 
                                  labels = c("Building with one apartment", 
                                             "Building with two apartments", 
                                             "Building with three or more apartments"))) |>
  dplyr::rename("Building class" = building_class) |>
  spread(retrofitted, `Mean HEPI (kWh/m2/year)`) |>
  rename(`Retrofitted` = "Yes", `Not retrofitted` = "No") |>
  mutate(`Difference in %` = 100 * (`Retrofitted` - `Not retrofitted`) / `Not retrofitted`) |>
  knitr::kable(booktabs = TRUE, digits = 1) 

retrofit_mean_hepi
```

```{r}
#| label: fig-retrofitting-investment-costs
#| fig-cap: "Correlation between retrofit investment costs and HEPI"

retrofit_investment_plot <- energy_modelling |>
  filter(!is.na(retrofit_investment_costs) & retrofit_investment_costs > 0) |>
  ggplot((aes(retrofit_investment_costs, hepi))) +
           geom_point()

retrofit_investment_plot
```

```{r}
#| label: fig-retrofitting
#| fig-cap: "Construction year per municipality"

retrofit_plot <- energy_modelling |>
  filter(retrofitted == "Yes") |>
  arrange(desc(egid)) |>
  distinct(egid, survey_year, retrofit_year_completion, energy_consumed_measured_mwh) |>
  pivot_wider(names_from = survey_year, names_prefix = "energy_consumed_measured_mwh_", values_from = energy_consumed_measured_mwh) 

retrofit_plot_2016_2018 <- retrofit_plot |>
  ggplot((aes(energy_consumed_measured_mwh_2018, energy_consumed_measured_mwh_2016))) +
           geom_point() +
  xlim(0,200) +
  ylim(0,200) +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = 'HEPI of buildings retrofitted between 2016 and 2018')

retrofit_plot_2018_2020 <- retrofit_plot |>
  ggplot((aes(energy_consumed_measured_mwh_2018, energy_consumed_measured_mwh_2020))) +
           geom_point() +
  xlim(0,200) +
  ylim(0,200) +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = 'HEPI of buildings retrofitted between 2018 and 2020')

retrofit_plot_2016_2020 <- retrofit_plot |>
  ggplot((aes(energy_consumed_measured_mwh_2016, energy_consumed_measured_mwh_2020))) +
           geom_point() +
  xlim(0,200) +
  ylim(0,200) +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = 'HEPI of buildings retrofitted between 2018 and 2020')
```

```{r}
retrofit_plot_2016_2018
retrofit_plot_2018_2020
retrofit_plot_2016_2020
```

### Municipalities

```{r}
#| label: indicators-municipalities
#| echo: false

#> Calculations of numbers and indicators for easier use as inline code in text.
n_municipalities <- n_distinct(energy_modelling$municipality_code)
```

In this section, we will have a closer look at the characteristics of the `r n_municipalities` that are represented in the data. Exploring characteristics and potential differences on the municipal level can sustain the modelling process by providing insights into the distribution and variability of energy consumption patterns among different municipalities.

```{r}
#| label: tbl-summary-municipality
#| tbl-cap: "Summary statistics per municipality"
#| tbl-cap-location: bottom

#> Filter to relevant buildings for modelling

summary_table <- energy_modelling |>
  group_by(entrance_municipality_name, building_class) |>
  summarise(n = n()) |>
  pivot_wider(names_from = building_class, 
              values_from = n)

summary_table
```

```{r}
#| label: boxplot-municipality
#| include: false

#> basic function
plot_boxplot <- function(df, y_axis) {
  y_axis <- sym(y_axis)
  df |>
    mutate(class = fct_reorder(entrance_municipality_name, !!y_axis, .fun='median')) |>
    ggplot( aes(x=reorder(entrance_municipality_name, !!y_axis), y=!!y_axis, fill=entrance_municipality_name)) + 
      geom_boxplot() +
      xlab("class") +
      theme(legend.position="none") +
      xlab("") +
      labs(paste0(y_axis," per municipality")) +
      coord_flip()
}
#> Creating boxplots for different input variables
boxplot_construction_year <- plot_boxplot(energy_modelling, "construction_year")
boxplot_building_area_m2 <- plot_boxplot(energy_modelling, "building_area_m2")
boxplot_num_residents <- plot_boxplot(energy_modelling, "num_residents")
boxplot_hepi <- plot_boxplot(energy_modelling, "hepi")

```

```{r}
#| label: fig-municipality-hepi
#| fig-cap: "HEPI per municipality"
#| include: false

ggplotly(boxplot_hepi)

```

```{r}
#| label: fig-municipality-construction-year
#| fig-cap: "Construction year per municipality"

ggplotly(boxplot_construction_year) 

```

### Ridgeline plots

```{r}
#| label: fig-hepi-municipality
#| fig-cap: "HEPI density plot per municipality"

energy_modelling |>
  filter(building_class == 1110) |>
  ggplot(aes(x = hepi, y = entrance_municipality_name, fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 5, rel_min_height = 0.02) +
  scale_fill_viridis(name = "HEPI (kWh/m2/year)", option = "C") +
  labs(title = 'HEPI per municipality') +
  xlim(0,300) +
   theme(axis.title.x=element_blank())
```

## Modelling
### Splitting the data

```{r}
#| label: train-test-split
#| include: false

#> select variables for modelling

         
#> Set the random number stream using `set.seed()` so that the results can be 
#> reproduced later. 
set.seed(501)

#> Save the split information for an 80/20 split of the data
energy_split <- initial_split(energy_modelling, 
                              prop = 0.01, 
                              strata = hec)
energy_split


#> Creating train and test set
energy_train <- training(energy_split)
energy_test  <-  testing(energy_split)


#> Creating a validation set (20% of training data)
set.seed(234)
energy_validation <- validation_split(energy_train,
                                      prop = 0.90,
                                      strata = hec)

dim(energy_train)
dim(energy_validation)

# creating folds for cross validation
set.seed(1502)
energy_folds <- 
   vfold_cv(energy_train, strata = hec, repeats = 3)
```

### Preprocessing Recipes

In the preprocessing, we apply the following:

-   Remove multicollinearity: Predictor variables that have a correlation above +/- 0.8 will me removed
-   Normalize numeric variables
-   Log-transform numeric variables
-   Create dummy variables from all nominal variables
   
```{r}
#| label: preprocessing-base-recipe-hec
#| include: false

# Specify basic recipe (applicable to all models)
recipe_base_hec <- recipe(hec ~ .,
                          data = energy_train) |> 
   step_select(egid, 
         hepi, 
         hec, 
         survey_year, 
         num_residents, 
         num_floors, 
         building_area_m2, 
         gross_floor_area_m2, 
         heated_area_m2, 
         year_of_installation, 
         efficiency_of_installation, 
         energy_usage_of_installation, 
         solar_system, 
         solar_system_area_m2, 
         solar_system_usage, 
         solar_system_area_m2, 
         photovoltaic_system, 
         photovoltaic_system_power_kw, 
         municipality_code, 
         building_class, 
         construction_year, 
         coordinate_e, 
         coordinate_n, 
         num_dwellings, 
         meters_above_sealevel, 
         energy_production_solar_mwh, 
         energy_consumed_hot_water_mwh,
         retrofitted,
         retrofit_investment_costs,
         hdd,
         hepi_pred_current_method,
         hec_pred_current_method) |>
  # remove variables from predictors that are only used for info or later comparison 
  update_role(egid, hepi, hepi_pred_current_method, hec_pred_current_method, new_role = "id") |>
  # Create dummy variables for nominal predictors
  step_dummy(all_nominal()) |>
  # Remove variables that contain only one value
  step_zv(all_numeric_predictors()) |>
  # Impute means for variables with NA
  step_impute_mean(num_residents) |>
  step_impute_mean(year_of_installation)

recipe_base_hec
```

```{r}
#| label: preprocessing-base-recipe-hec-norm
#| include: false

#> Some models (notably neural networks, KNN, and support vector machines) require predictors that have been centered and scaled, so some model workflows will require recipes with these preprocessing steps. For other models, a traditional response surface design model expansion (i.e., quadratic and two-way interactions) is a good idea.
#> https://www.tmwr.org/workflow-sets.html

recipe_base_hec_norm <- recipe_base_hec |>
  step_normalize(all_numeric_predictors())

recipe_base_hec_norm
```

```{r}
#| label: preprocessing-base-recipe-hec-poly
#| include: false

# Add polynomials and interaction
recipe_base_hec_poly <- 
   recipe_base_hec_norm %>% 
   step_poly(all_numeric_predictors()) %>% 
   step_interact(~ all_numeric_predictors():all_numeric_predictors())

recipe_base_hec_poly
```


### Defining models

```{r}
#| label: model-definition
#| include: false

# MLR model
lm_basic_spec <- 
   linear_reg() |> 
   set_engine("lm")

# MLR glmnet spec
lm_glmnet_spec <- 
   linear_reg(penalty = tune(), mixture = tune()) %>% 
   set_engine("glmnet")

# decision tree
dt_spec <- 
   decision_tree() |> 
   set_engine("rpart") |> 
   set_mode("regression")

# k nearest neighbour
knn_spec <-
   nearest_neighbor(neighbors = tune(), dist_power = tune(), weight_func = tune()) %>%
   set_engine("kknn") %>%
   set_mode("regression")

# random forest
rf_spec <- rand_forest() |>
  set_engine("ranger") |>
  set_mode("regression")

# xgboost
xgb_spec <- 
   boost_tree(tree_depth = tune(), learn_rate = tune(), loss_reduction = tune(), 
              min_n = tune(), sample_size = tune(), trees = tune()) %>% 
   set_engine("xgboost") %>% 
   set_mode("regression")

# svm radial
svm_r_spec <-
   svm_rbf(cost = tune(), rbf_sigma = tune()) %>%
   set_engine("kernlab") %>%
   set_mode("regression")

# svm poly
svm_p_spec <-
 svm_poly(cost = tune(), degree = tune()) %>%
 set_engine("kernlab") %>%
 set_mode("regression")


# articifial neural net
nnet_spec <- 
   mlp(hidden_units = tune(), penalty = tune(), epochs = tune()) %>% 
   set_engine("nnet", MaxNWts = 2600) %>% 
   set_mode("regression")

# auto ml
auto_ml_spec <- auto_ml() |>
  set_engine("h2o") |>
  set_mode("regression") 
  
```

```{r}
#| label: model-definition-tmwr
#| include: false

#> https://www.tmwr.org/workflow-sets.html#creating-the-workflow-set

# mars_spec <- 
#    mars(prod_degree = tune()) %>%  #<- use GCV to choose terms
#    set_engine("earth") %>% 
#    set_mode("regression")
# 
# svm_r_spec <- 
#    svm_rbf(cost = tune(), rbf_sigma = tune()) %>% 
#    set_engine("kernlab") %>% 
#    set_mode("regression")
# 
# svm_p_spec <- 
#    svm_poly(cost = tune(), degree = tune()) %>% 
#    set_engine("kernlab") %>% 
#    set_mode("regression")
# 
# knn_spec <- 
#    nearest_neighbor(neighbors = tune(), dist_power = tune(), weight_func = tune()) %>% 
#    set_engine("kknn") %>% 
#    set_mode("regression")
# 
# cart_spec <- 
#    decision_tree(cost_complexity = tune(), min_n = tune()) %>% 
#    set_engine("rpart") %>% 
#    set_mode("regression")
# 
# bag_cart_spec <- 
#    bag_tree() %>% 
#    set_engine("rpart", times = 50L) %>% 
#    set_mode("regression")
# 
# rf_spec <- 
#    rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
#    set_engine("ranger") %>% 
#    set_mode("regression")
# 
# 
# 
# cubist_spec <- 
#    cubist_rules(committees = tune(), neighbors = tune()) %>% 
#    set_engine("Cubist") 
```

### Define workflows


#### Workflow for nonlinear models
```{r}
#| label: workflow-definition-normalized
#| include: false

#> this workflow is for nonlinear models that require the predictors to be in the same units

normalized <- 
   workflow_set(
      preproc = list(normalized = recipe_base_hec_norm), 
      models = list(neural_network = nnet_spec, 
                    SVM_radial = svm_r_spec, 
                    SVM_poly = svm_p_spec, 
                    KNN = knn_spec)
   )

normalized
```
#### Workflow with only basic preprocessing

```{r}
#| label: workflow-definition-basic
#| include: false

#> this workflow is for models that do not need normalizing

no_pre_proc <- 
   workflow_set(
      preproc = list(simple = recipe_base_hec), 
      models = list(lm_basic = lm_basic_spec, 
                    random_forest = rf_spec, 
                    xgboost = xgb_spec
                    )
   )

no_pre_proc
```

#### Models with non linear terms and interactions

```{r}
#| label: workflow-definition-with-features
#| include: false

with_features <- 
   workflow_set(
      preproc = list(full_quad = recipe_base_hec_poly), 
      models = list(lm_glmnet = lm_glmnet_spec, KNN = knn_spec)
   )
```

#### Combine all workflows

```{r}
#| label: combine-workflows
#| include: false

all_workflows <- 
   bind_rows(no_pre_proc, normalized) %>% 
   # Make the workflow ID's a little more simple: 
   mutate(wflow_id = gsub("(simple_)|(normalized_)", "", wflow_id))

all_workflows
```


### Tuning

```{r}
#| label: hyperparameter-tuning
#| include: false

#> Almost all of the members of all_workflows contain tuning parameters. To evaluate their performance, we can use the standard tuning or resampling functions (e.g., tune_grid() and so on). The workflow_map() function will apply the same function to all of the workflows in the set; the default is tune_grid()


grid_ctrl <-
   control_grid(
      save_pred = TRUE,
      parallel_over = "everything",
      save_workflow = FALSE
   )

grid_results <-
   all_workflows %>%
   workflow_map(
      seed = 1503,
      resamples = energy_validation,
      grid = 25,
      control = grid_ctrl
   )

```

```{r}
#| label: grid-results
#| include: false

# without knn
grid_results_subset <- grid_results %>%
   subset(wflow_id != "KNN")

autoplot(
   grid_results_subset,
   rank_metric = "rmse",  # <- how to order models
   metric = "rmse",       # <- which metric to visualize
   select_best = TRUE     # <- one point per workflow
) +
   geom_text(aes(y = mean - 1/2, label = wflow_id), angle = 90, hjust = 1) +
   #lims(y = c(3.5, 9.5)) +
   theme(legend.position = "none") +
   ylim(18000, 28000) +
   ggtitle("Benchmark") +
   ylab("RMSE")
```

#### Improve effiecency: Racing approach
```{r}
#| label: race-results
#| include: false

race_ctrl <-
   control_race(
      save_pred = TRUE,
      parallel_over = "everything",
      save_workflow = TRUE
   )

race_results <-
   all_workflows %>%
   workflow_map(
      "tune_race_anova",
      seed = 1503,
      resamples = energy_folds,
      grid = 25,
      control = race_ctrl
   )

race_results
```

```{r}
#| label: race-autoplot
#| include: true

autoplot(
   race_results,
   rank_metric = "rmse",  
   metric = "rmse",       
   select_best = TRUE    
) +
   geom_text(aes(y = mean - 1/2, label = wflow_id), angle = 90, hjust = 1) +
   lims(y = c(3.0, 9.5)) +
   theme(legend.position = "none")
```

#### Compare grid search and race approach

```{r}
#| label: compare-grid-race
#| include: true
matched_results <- 
   rank_results(race_results, select_best = TRUE) %>% 
   select(wflow_id, .metric, race = mean, config_race = .config) %>% 
   inner_join(
      rank_results(grid_results, select_best = TRUE) %>% 
         select(wflow_id, .metric, complete = mean, 
                config_complete = .config, model),
      by = c("wflow_id", ".metric"),
   ) %>%  
   filter(.metric == "rmse")

matched_results %>% 
   ggplot(aes(x = complete, y = race)) + 
   geom_abline(lty = 3) + 
   geom_point() + 
   geom_text_repel(aes(label = model)) +
   coord_obs_pred() + 
   labs(x = "Complete Grid RMSE", y = "Racing RMSE") 
```

### Training

```{r}
#| label: model-fit-basic
#| include: false


doParallel::registerDoParallel()

fit_basic <- workflow_map(wf_set_basic,
                          "fit_resamples", 
                          resamples = energy_folds,
                          metrics = metric_set(rmse, mape, mae, rsq),
                          verbose = TRUE,
                          seed = 22)

fit_basic
```

### Model evaluation {#sec-model-evaluation}

```{r}
#| label: metrics-current-approach
#| include: false

energy_train_current_method <- energy_train |>
  dplyr::inner_join(energy_modelling, by = c("egid"), suffix = c("", "_y")) |>
  dplyr::select(egid, survey_year, hec, hepi, hepi_pred_current_method, hec_pred_current_method, heated_area_m2)

custom_metrics <- metric_set(rmse, mape, mae, rsq)

energy_train_metrics_current_method <- energy_train_current_method |>
  custom_metrics(hec, hec_pred_current_method)

energy_train_metrics_current_method


results_train_hec_current_method <- energy_train |> 
   ggplot(aes(x = hec, y = hec_pred_current_method)) + 
   geom_abline(color = "gray50", lty = 2) + 
   geom_point(alpha = 0.5) + 
   coord_obs_pred() + 
   labs(x = "Observed HEC", y = "Predicted HEC")

results_train_hec_current_method
```




```{r}
#| label: fig-model-benchmarking
#| fig-cap: "Benchmarking of models"
fit_basic |>
   rena
autoplot(fit_basic)
```

```{r}
#| label: fig-model-benchmarking-ranking-rmse
#| fig-cap: "Benchmarking of models: RMSE ranking"

fit_basic_metrics <- collect_metrics(fit_basic)

workflowsets::rank_results(fit_basic, rank_metric = "mape", select_best = TRUE)

fit_basic_metrics
```

```{r}
autoplot(fit_basic)
```

### Final model {#sec-final-model}

```{r}
best_model_xgboost <- 
   grid_results %>% 
   extract_workflow_set_result("xgboost") %>% 
   select_best(metric = "rmse")

best_model_xgboost

test_results_xgboost <- 
   grid_results %>% 
   extract_workflow("xgboost") %>% 
   finalize_workflow(best_model_xgboost) %>% 
   last_fit(split = energy_split)


```

```{r}
collect_metrics(test_results_xgboost)

test_results_xgboost %>% 
   collect_predictions() %>% 
   ggplot(aes(x = hec, y = .pred)) + 
   geom_abline(color = "gray50", lty = 2) + 
   geom_point(alpha = 0.5) + 
   coord_obs_pred() + 
   labs(x = "observed", y = "predicted")

```

### Save models

```{r}
#| label: save-models
#| include: false
xgboost_wf <- extract_workflow(test_results_xgboost)

readr::write_rds(test_results_xgboost, "models/xgboost_wf")
```

